---
title: "Extra grounds inventory and extraction metadata"
author: "Anna Goodman"
date: "2026-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# read main tab from EM Metadata entry spreadsheet
```{r}
metaMain = read_csv("EM Metadata entry, SEKI 2024 - AG_direct_edit_22April2026.csv") %>% # load in main root metadata sheet
  mutate(Extraction_ID = as.numeric(Extraction_ID)) %>%
  filter(!is.na(Extraction_ID))

head(metaMain) # look at it

```

# read in tab from EM Metadata entry spreadsheet with info on how much extra ground roots we have for each sample
```{r}
xtra_grid = read_csv("EM Metadata entry, SEKI 2024 - Xtra grounds volume.csv") # load in volumes of extra roots

head(xtra_grid) # look at it

xtra_grid = xtra_grid[, 1:7] %>% # trim to exclude notes columns
  filter(`Box dated` != "totals:" & `Box dated` != "percents:")

head(xtra_grid) # look at it again
max(xtra_grid$Extraction_ID) # look at it again

```

# make a single column in xtra_grid that says how much extra root grounds are left. A is least, E is most
```{r}
xtra_grid = mutate(xtra_grid, volXtra = "placeholder")
for (i in 1:length(xtra_grid$Extraction_ID)) {
  if (is.na(xtra_grid$A[i]) == FALSE) {
    xtra_grid$volXtra[i] = "A"
  } else if (is.na(xtra_grid$B[i]) == FALSE) {
    xtra_grid$volXtra[i] = "B"
  } else if (is.na(xtra_grid$C[i]) == FALSE) {
    xtra_grid$volXtra[i] = "C"
  } else if (is.na(xtra_grid$D[i]) == FALSE) {
    xtra_grid$volXtra[i] = "D"
  } else if (is.na(xtra_grid$E[i]) == FALSE) {
    xtra_grid$volXtra[i] = "E"
  }
}
head(xtra_grid)
```

# view which samples have most extra root grounds 
```{r}
xtra_grid = xtra_grid[rev(order(xtra_grid$Extraction_ID)),] # reorder dataframe by extraction id, low to high
xtra_grid = xtra_grid[rev(order(xtra_grid$volXtra)),] # reorder dataframe by volXtra, with most volume first
head(xtra_grid)

xtra_table = xtra_grid %>%
  select(`Box dated`, Extraction_ID, volXtra) %>% # move useful columns to a simpler dataframe
  rename(xtra_box_label = `Box dated`)
head(xtra_table)
```

# add info: extraction date, qubit results, dilutions where availible
## option 1: from dna dilution metadata
```{r}
# note which ones have already been extracted

done_extractions = read_csv("../dilutions/Results/DNA_dilutions_all_striptubes.csv")

done_extractions = select (done_extractions, !c(`...1`)) #  `...1` may appear as a column name as an artifact

# update remaining_original_ul so it's what it will be once we do the 2ng/ul dilutions
done_extractions = done_extractions %>%
  mutate(remaining_original_ul = remaining_original_ul - round_original) %>%
  rename(dilution_vol_ul = final_vol_ul, dilution_ng_per_ul = final_ng_per_ul)

extraction_metadata = metaMain %>%
  left_join(xtra_table, join_by("Extraction_ID")) %>%
  left_join(done_extractions, join_by("Extraction_ID" == "extraction_id")) 
head(extraction_metadata)
nrow(extraction_metadata) # too many rows. Checked it out, it's because xtra_table has a few duplicate rows; 

extraction_metadata = extraction_metadata %>%
  filter(!(Extraction_ID == 39 & volXtra == "A")) %>% # 39 reported as having A extra and D extra, in an august and december box. Removing december box with A extra, because the grinding happened in august 2025.
  filter(!(Extraction_ID == 47 & volXtra == "B")) %>% #47 appears twice in xtra_table, once with box dated nov 2026. It was extracted in 2026, so I'm keeping the other one
    filter(!(Extraction_ID == 63 & volXtra == "E")) # 63 appears twice in xtra_table, box dates equivalent. Taking out entry with more volume to be conservative

nrow(extraction_metadata)
```
```{r}
write.csv(extraction_metadata, "Results/extraction_metadata_w_dilutions.csv")
```

