---
title: "august2026_audit"
output: html_document
---

```{r}
library(tidyverse)
library(googledrive)
```

```{r}
mainMetadata_dribble = drive_get("EM Metadata entry, SEKI 2024", shared_drive = 'Bogar Lab')

mainMetadata_dribble
```
```{r}
local_path <- paste0("Inputs/EM Metadata entry, SEKI 2024_downloaded-", format(Sys.Date(), "%Y%m%d"), ".csv") #define what this local file name will be, with today's date

drive_download(mainMetadata_dribble, type = "csv", path = local_path, overwrite = TRUE) # overwrite = TRUE means that if you download this file twice on the same day, the first version will be overwritten by the second. If you download this file again on a different day (by running this whole code chunk), it will not be overwritten, because the date will have been updated and the local path will be different
```
```{r}
mainMetadata_dribble = drive_get("EM Metadata entry, SEKI 2024", shared_drive = 'Bogar Lab')

mainMetadata_dribble
```
```{r}
xtra_roots_dribble = drive_get("xtra_root_grounds_volume", shared_drive = 'Bogar Lab')

xtra_roots_dribble

local_path <- paste0("Inputs/xtra_root_grounds_volume_downloaded", format(Sys.Date(), "%Y%m%d"), ".csv") 

drive_download(xtra_roots_dribble, type = "csv", path = local_path, overwrite = TRUE)
```
```{r}

xtra_rootscsv = read.csv("Inputs/xtra_root_grounds_volume_downloaded20260812.csv") 


# metadatacsv = read.csv("Inputs/EM Metadata entry, SEKI 2024_downloaded-20260812.csv") TEMPORARY: anna re-running with the local, "AG direct edit" version on her computer

metadatacsv = read.csv("Inputs/EM Metadata entry, SEKI 2024 - AG_direct_edit_22April2026.csv")
```


```{r}
xtra_grid = mutate(xtra_rootscsv, volXtra = "placeholder") # made a column called volXtra

head(xtra_grid)

for (i in 1:length(xtra_grid$Extraction_ID)) {
  if ((xtra_grid$A[i]) == "x") {
    xtra_grid$volXtra[i] = "A"
  } else if ((xtra_grid$B[i]) == "x") {
    xtra_grid$volXtra[i] = "B"
  } else if ((xtra_grid$C[i]) == "x") {
    xtra_grid$volXtra[i] = "C"
  } else if ((xtra_grid$D[i]) == "x") {
    xtra_grid$volXtra[i] = "D"
  } else if ((xtra_grid$E[i]) == "x") {
    xtra_grid$volXtra[i] = "E"
  }
}
head(xtra_grid)

xtra_grid = xtra_grid %>% # make a copy of xtra_grid, with the following changes:
  select(c(Box.dated, Extraction_ID, volXtra)) # only keep the columns with these names!
      # the c() makes a "vector" (basically, a list) of all the columns we're keeping. you wouldn't need it if you were just keeping one column.

head(xtra_grid)

xtra_grid = xtra_grid %>%
  filter(Box.dated != "totals: ") %>%
  filter(Box.dated != "percents:")

```
Join the main metadata table with the extra root volumes
```{r}
xtra_grid$Extraction_ID <- as.character(xtra_grid$Extraction_ID) # 

Joint_table = left_join(metadatacsv,xtra_grid,by = "Extraction_ID")
```

check for joining errors:
```{r}
quality_check = Joint_table %>% # make a copy of Joint_table, with the following modifications:
  select(c(Extraction_ID, date_ground_english, volXtra, Box.dated)) # just look at these columns 

# View(quality_check) # ok this looks good, but I think we lost one of the two 39s in the xtra grounds sheet and I don't know which one. re-trying with full_join
```

resolving join errors
```{r}
# There were two Extraction_ID 39s in the extra grounds sheet; based on dates, one of them is actually 59 (which is otherwise absent from that sheet.) Verified by looking in freezer on 11 August 2026, EF and AG.
  # So we know that #59 has "A" volXtra and is in Box.dated 5 December, 2025

Joint_table = Joint_table %>%
  filter(!(Extraction_ID == "39" & Box.dated == "5 December 2025")) %>%
  mutate(volXtra = case_when(
    Extraction_ID == "59" ~ "A",
    .default = volXtra
  )) %>%
    mutate(Box.dated = case_when(
    Extraction_ID == "59" ~ "5 December 2025",
    .default = Box.dated
  ))
  
```

Remove all columns where extraction id isn't numeric, make all extraciton ids numeric
```{r}
Joint_table = Joint_table %>%
  mutate(Extraction_ID = as.numeric(Extraction_ID)) %>%
  filter(!is.na(Extraction_ID))
```
Keep looking for joining errors
```{r}
Joint_table = Joint_table %>%
  arrange(Extraction_ID)

98 %in% Joint_table$Extraction_ID #there's a 98 in the joined version
"98" %in% metadatacsv$Extraction_ID #it was also in the metadatacsv
98 %in% Joint_table$xtra_rootscsv #it was not in the xtrarootscsv

anyDuplicated(Joint_table$Extraction_ID) #should give me the index (row number) of a duplicated value

Joint_table$Extraction_ID[anyDuplicated(Joint_table$Extraction_ID)]
```

check for joining errors:
```{r}
quality_check = Joint_table %>% # make a copy of Joint_table, with the following modifications:
  select(c(Extraction_ID, date_ground_english, volXtra, Box.dated)) # just look at these columns 

# View(quality_check) # ok this looks good, but I think we lost one of the two 39s in the xtra grounds sheet and I don't know which one. re-trying with full_join
```

resolving join errors
```{r}
# There were two Extraction_ID 47s in the extra grounds sheet; based on dates, one of them is actually 98 (which is otherwise absent from that sheet.) This is related to the 47->98 labeling switch, documented on Anna's OneNote and also in the AG direct edit version of the main metadata sheet. 
  # So we know that #98 has "D" volXtra and has "I forgot to note" in Box.dated

Joint_table = Joint_table %>%
  filter(!(Extraction_ID == "47" & Box.dated == "I forgot to note")) %>%
  mutate(volXtra = case_when(
    Extraction_ID == "98" ~ "D",
    .default = volXtra
  )) %>%
    mutate(Box.dated = case_when(
    Extraction_ID == "98" ~ "I forgot to note",
    .default = Box.dated
  ))
  
```

Keep looking for joining errors
```{r}
anyDuplicated(Joint_table$Extraction_ID) #should give me the index (row number) of a duplicated value

Joint_table$Extraction_ID[anyDuplicated(Joint_table$Extraction_ID)] # should give me the value of Extraction_ID in that row
```

resolving join errors
```{r}
# There were two Extraction_ID 63s in the extra grounds sheet; based on dates, one of them is actually 65 (which is otherwise absent from that sheet.) Verified by looking in freezer on 11 August 2026, EF and AG.
  # So we know that #65 has "B" volXtra and has "5 Decemeber 2025" in Box.dated

Joint_table = Joint_table %>%
  filter(!(Extraction_ID == "63" & volXtra == "B")) %>%
  mutate(volXtra = case_when(
    Extraction_ID == "65" ~ "B",
    .default = volXtra
  )) %>%
    mutate(Box.dated = case_when(
    Extraction_ID == "65" ~ "5 December 2025",
    .default = Box.dated
  ))
  
```

Keep looking for joining errors
```{r}
anyDuplicated(Joint_table$Extraction_ID) #should give me the index (row number) of a duplicated value

Joint_table$Extraction_ID[anyDuplicated(Joint_table$Extraction_ID)] # should give me the value of Extraction_ID in that row

quality_check = Joint_table %>% # make a copy of Joint_table, with the following modifications:
  select(c(Extraction_ID, date_ground_english, volXtra, Box.dated)) # just look at these columns 

```

```{r}
dilutions_dataframe = read_csv("../../dilutions/Results/DNA_dilutions_all_striptubes.csv") %>%
  rename(Extraction_ID = extraction_id)
```

Join this with the dilutions I planned 
```{r}
Joint_w_dil = left_join(Joint_table, dilutions_dataframe, by = "Extraction_ID")

quality_check = Joint_w_dil %>% # make a copy of Joint_w_dil, with the following modifications:
  select(c(Extraction_ID, date_ground_english, volXtra, remaining_original_ul, round_original)) # just look at these columns; 
          # "remaining_original_ul" was how much DNA extract we had before the dilutions, 
          # "round_original" was how much DNA extract I planned to use for the dilutions
```

Ok, which ones have already been diluted?
We need to find 
- which extraction IDs Anna Diluted for LPA and LPBoot
- which extraction IDs Evan diluted for LPBoot and LPC, LPD

Some of this info exists on the drive in a library prep progress tracker document!
```{r}
LP_progress_dribble = drive_get("SEKI_Libraryprep_Progress", shared_drive = 'Bogar Lab')

LP_progress_dribble
```
```{r}
local_path <- paste0("Inputs/SEKI_Libraryprep_Progress_downloaded-", format(Sys.Date(), "%Y%m%d"), ".csv") #define what this local file name will be, with today's date

drive_download(LP_progress_dribble, type = "csv", path = local_path, overwrite = TRUE) # overwrite = TRUE means that if you download this file twice on the same day, the first version will be overwritten by the second. If you download this file again on a different day (by running this whole code chunk), it will not be overwritten, because the date will have been updated and the local path will be different

LP_progress = read_csv(local_path)
```

Which extraction IDs have been diluted, and how many times? And how much of the original volume is left now?

written w google gemini
```{r}
LP_summary <- LP_progress %>%
  group_by(Extraction_ID) %>%
  summarize(
    dilutions_completed = sum(!is.na(Date_DNA_Dilution)),
    LP_sucessful = case_when(
      any(`Sucsesfull-prep_(Y/N)` == "Y", na.rm = TRUE) ~ "Y",
      all(`Sucsesfull-prep_(Y/N)` == "N", na.rm = TRUE) ~ "N",
      .default = "TBD"
    ),
    
    # Concatenate original metadata, but EXCLUDE the new summary columns
    across(
      -c(`Sucsesfull-prep_(Y/N)`, dilutions_completed, LP_sucessful), 
      ~ paste(na.omit(unique(as.character(.x))), collapse = "; ")
    ),
    .groups = "drop"
  ) %>%
  # suppressWarnings prevents the console clutter from metadata rows 
  # like "# of samples completed" being turned into NA
  mutate(Extraction_ID = suppressWarnings(readr::parse_number(as.character(Extraction_ID)))) %>%
  filter(!is.na(Extraction_ID))


Joint_w_dil_progress <- Joint_w_dil %>% 
  select(-any_of(c("Date_DNA_Dilution", "LP_sucessful", "dilutions_completed"))) %>%
  mutate(
    Extraction_ID = suppressWarnings(readr::parse_number(as.character(Extraction_ID))),
    round_original = as.numeric(round_original),
    remaining_original_ul = as.numeric(remaining_original_ul)
  ) %>%
  left_join(
    select(LP_summary, -any_of(c("round_original", "remaining_original_ul"))), 
    by = "Extraction_ID"
  ) %>%
  mutate(
    ul_used_for_dilutions = dilutions_completed * round_original,
    remaining_original_ul = remaining_original_ul - ul_used_for_dilutions
  )

```


```{r}
Extra_inventory = Joint_w_dil_progress %>%
  select(c(Extraction_ID, volXtra, remaining_original_ul, dilutions_completed, LP_sucessful, Prep_ID, Date_DNA_Dilution, original_ng_per_ul)) %>%
  mutate(ng_DNA_undiluted = original_ng_per_ul * remaining_original_ul)
  arrange(volXtra) %>%
  arrange(remaining_original_ul)

nonrepeatable_or_thin_ice = Extra_inventory %>%
  filter(remaining_original_ul < 30) %>% # arbitrary "not a lot" volume
  filter(volXtra == "A") # least amount of extra root grounds

re_extract = Extra_inventory %>%
  filter(remaining_original_ul < 15) %>% # arbitrary "not a lot" volume
  filter(volXtra != "A") %>% # enough root grounds for a re-extraction
  filter(LP_sucessful == "N") %>% # library prep attempts thus far have failed
  filter(Date_DNA_Dilution == 20260728) # were in the batch of contaminated dilutions

re_dilute_from_original_extract = Extra_inventory %>%
  filter(remaining_original_ul > 15) %>% # arbitrary "enough" volume; could try to LP at least 10 times even if original concentration minimal
  filter(LP_sucessful == "N") %>% # library prep attempts thus far have failed
  filter(Date_DNA_Dilution == 20260728) # were in the batch of contaminated dilutions

plenty_of_extra = Extra_inventory %>%
  filter(remaining_original_ul > 50) %>% # arbitrary "lots" volume
  filter(LP_sucessful != "Y") %>% # LPs so far failed or not attempted
  filter(volXtra != "A") %>% # not almost no roots leftover
  filter(volXtra != "B") # also not just a little bit of roots left over
```

Export these inventories
```{r}
date_suffix <- paste0(format(Sys.Date(), "%Y%m%d"), ".csv") #define what this local file name will be, with today's date

write.csv(nonrepeatable_or_thin_ice, paste0("Results/nonrepeatable_or_thin_ice_", date_suffix))

write.csv(re_extract, paste0("Results/to_re_extract", date_suffix))

write.csv(re_dilute_from_original_extract, paste0("Results/to_re_dilute_from_original_extract_", date_suffix))

write.csv(plenty_of_extra, paste0("Results/plenty_of_extra_", date_suffix))

```



