---
title: "DNA_dilutions_first24"
author: "Anna Goodman"
date: "2026-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
q_data = read_csv("Qubit Results - raw_sample_results.csv") # read in qubit data
# 1 row = 1 measurement. Some extractions measured multiple times, where first measurement read out of range (OOR)

head(q_data)
```

```{r}
q_data = q_data %>% mutate(ng_per_ul = as.numeric(ng_per_ul)) # coerce the ng_per_ul column of this dataframe to be numeric; this will make the numbers class numeric and the "HIGH" or "LOW" entries (for OORs) NAs

is.numeric(q_data$ng_per_ul) #it worked
q_data$ng_per_ul # confirmed HIGH and LOW are now NA
```

```{r}
# make a dataframe without OOR measurements
in_range_qs = filter(q_data, !is.na(q_data$ng_per_ul)) # should remove non-numeric concentration results

head(in_range_qs) # not empty
nrow(in_range_qs) # 24 rows: good, one for each extraction
```
```{r}
# ok so now we need to count up all the ul consumed from each original extract
  ## from q_data
# and calculate how much is left of each extract 
  ## in in_range_qs

for (i in 1:length(in_range_qs$extraction_id)) { # for each original extraction,
  total_ul_consumed = 0 # initialize var to store total ul used for qubitting
  for (j in 1:length(q_data$extraction_id)) { # then look through each qubit measuremnt
    
    if (q_data$extraction_id[j] == in_range_qs$extraction_id[i]) { # if that qubit measurement was of this original extraction,
      total_ul_consumed = total_ul_consumed + q_data$ul_original_consumed[j] # add the # of ul consumed to make that quibit measurement to the total consumed from this original extraction
    } else {}
  }
  in_range_qs$ul_original_consumed[i] = total_ul_consumed # now store the total ul consumed for all measurements of this original extraction in in_range_qs
}

head(in_range_qs)
```

```{r}
# extrapolate original extraction's concentration for cases where we had to measure a dilution on the qubit

# convert the "dilution factor" into a proportion (original extract volume)/(H2O volume)
proportion = c() # put proportion values in a vector, in matching order with the data frame
for (i in 1:length(in_range_qs$extraction_id)) {
  dilfac = in_range_qs$dilution_factor[i]
  if (dilfac == "original_extract") {
    proportion[i] = 1
  } else if (dilfac == "1:1") {
    proportion[i] = 0.5
  } else if (dilfac == "1:3") {
    proportion[i] = 0.25
  }
}

in_range_qs = mutate (in_range_qs, proportion_original = proportion) # make that vector a column in in_range_qs


concs = mutate (in_range_qs, extrapolated_original_ng_per_ul = (1/proportion_original)*ng_per_ul) # calculate original concentration from measured concentration of dilution

head(concs)
```




```{r}
concs = mutate(concs, remaining_original_ul = 65 - ul_original_consumed) # make new column with remaining ul of original extract. all extracts from this extraction day eluted in 75ul of elution buffer, -10ul for safety/EB that never made it out of the spin column

concs$remaining_original_ul # looks like that worked
```

# Now, we see what we need to do to dilute each extract to 2ng/ul 

c (concentration) = m (mass) / v (volume)

current state (1): we have a finite mass of DNA, in elution buffer, at a measured concentration (ng/uL)
    # c1 = qubit-measured concentration in ng/ul ("extrapolated_original_ng_per_ul")
    # m1 = total mass of DNA in usable volume of original extract
    # v1 = usable volume of original extract ("remaining_original_ul")
state 2: we want to use all the DNA in that usable volume of extract to make a 2ng/ul solution of unknown volume
    # c2 = 2ng/ul
    # m2 = m1; total mass of DNA in diluted extract (we're using all of the usable original)
    # v2 = total volume of diluted extract
* we need to find v2! and v2 will be made of v1+H2O, so volume of H2O needed will be v2-v1


c = m/v, so cv = m
m2 = m2, so c1v1 = c2v2, so (c1v1)/c2 = v2
and H2O needed = v2-v1

```{r}
# calculate total volume of diluted extract
v2 = function(c1, v1, c2) {
  v2 = (c1*v1)/c2
  return(v2)
}

concs = mutate(concs, # add 2 columns: final volume of diluted extract, volume of H2O to add
               final_vol_ul = v2(extrapolated_original_ng_per_ul, remaining_original_ul, 2),
               H2O_ul = final_vol_ul - remaining_original_ul)
head(concs)


```

```{r}
# some big numbers in there. want to use all 1.5 mL tubes for these dilutions
tube_fit_problem = filter(concs, concs$final_vol_ul >= 1450)
tube_fit_problem #51 is pushing it but will probably fit. #3 will not

concs = mutate(concs, ul_original_into_final = remaining_original_ul) # make a separate column for v1, showing what volume of original extract should go into the final dilution. set default as all remaining_original_ul

# chance the ul_original_into_final volume, just for #3.

# constrain v2 and find v1; (c2v2)/c1 = v2
v1 = function(c2, v2, c1) {
  v1 = (c2*v2)/c1
  return(v1)
}

index = which(concs$extraction_id == 3) # store the index of extraction ID 3
concs[index, "final_vol_ul"] = 1450 # set final vol of dilution

concs[index, "ul_original_into_final"] = v1(2, 
                                                         concs$final_vol_ul[index], # final vol set above
                                                         concs$extrapolated_original_ng_per_ul[index]) # replace the v1 value in concs for #3

concs = mutate(concs, # re-write h2o 
               H2O_ul = final_vol_ul - ul_original_into_final)
head(concs)
```








```{r}
# calculate final concentration to double-check

c2 = function(c1, v1, v2) {
  c2 = (c1*v1)/v2
  return(c2)
}

concs = mutate(concs, 
               final_ng_per_ul = c2(extrapolated_original_ng_per_ul, ul_original_into_final, final_vol_ul))

concs$final_ng_per_ul # all 2s!
```
```{r}
# tidy unnecessary columns, rename for clarity
DNA_dilutions_first24 = select(concs, !c(qubit_run_id, dilution_factor, ul_original_consumed, proportion_original, ng_per_ul))

DNA_dilutions_first24 = rename(DNA_dilutions_first24, ng_per_ul_original = extrapolated_original_ng_per_ul)
                               
```


```{r}
write.csv(DNA_dilutions_first24, "Results/DNA_dilutions_first24.csv")
```

# OR what if we did smaller volume dilutions, like into an 8-strip (200ul capacity? let's go for 180ul)
```{r}
# v2 = 180ul
# c2 = 2 ng/ul
# c1 is measured concentration

# find v1

smallTubes = select(DNA_dilutions_first24, !c())

```




```{r}
already_good = filter(concs, concs$final_vol_ul <= 180)


# constrain v2 and find v1; (c2v2)/c1 = v2
v1 = function(c2, v2, c1) {
  v1 = (c2*v2)/c1
  return(v1)
}

index = which(concs$extraction_id !in already_good$extraction_id) # indicies or 
concs[index, "final_vol_ul"] = 1450 # set final vol of dilution

concs[index, "ul_original_into_final"] = v1(2, 
                                                         concs$final_vol_ul[index], # final vol set above
                                                         concs$extrapolated_original_ng_per_ul[index]) # replace the v1 value in concs for #3

concs = mutate(concs, # re-write h2o 
               H2O_ul = final_vol_ul - ul_original_into_final)
head(concs)
```

