---
title: "DNA_dilutions_all"
author: "Anna Goodman"
date: "2026-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
q_data = read_csv("../Qubit Results - raw_sample_results.csv") # read in qubit data
# 1 row = 1 measurement. Some extractions measured multiple times, where first measurement read out of range (OOR)

head(q_data)
```

```{r}
# make a dataframe without OOR measurements
oor_strings = c("High", "Low")
in_range_qs = q_data %>%
  filter(!(ng_per_ul %in% oor_strings | ng_per_ml %in% oor_strings)) %>% # should remove all qubits with out of range results
  mutate(ng_per_ul = as.numeric(ng_per_ul),
         ng_per_ml = as.numeric(ng_per_ml)) # coerce all concentration data to numeric type; it was character before because of the OOR values



head(in_range_qs) # not empty
is.numeric(in_range_qs$ng_per_ul) #it worked
is.numeric(in_range_qs$ng_per_ml) #it worked
```



```{r}
# ok so now we need to count up all the ul consumed from each original extract
  ## from q_data
# and calculate how much is left of each extract 
  ## in in_range_qs

for (i in 1:length(in_range_qs$extraction_id)) { # for each original extraction,
  total_ul_consumed = 0 # initialize var to store total ul used for qubitting
  for (j in 1:length(q_data$extraction_id)) { # then look through each qubit measuremnt
    
    if (q_data$extraction_id[j] == in_range_qs$extraction_id[i]) { # if that qubit measurement was of this original extraction,
      total_ul_consumed = total_ul_consumed + q_data$ul_original_consumed[j] # add the # of ul consumed to make that quibit measurement to the total consumed from this original extraction
    } else {}
  }
  in_range_qs$ul_original_consumed[i] = total_ul_consumed # now store the total ul consumed for all measurements of this original extraction in in_range_qs
}

head(in_range_qs)
```

```{r}
qs_ng_ul = in_range_qs %>%
  mutate(
    ng_per_ul = case_when(
      !is.na(ng_per_ul) ~ ng_per_ul,
      is.na(ng_per_ul) ~ ng_per_ml/1000
                               )
         ) # for each row where ng_per_ul column is empty, grab data from ng_per_ml column, convert to ng/ul, and store in  ng_per_ul column

head(qs_ng_ul)
```




```{r}
# extrapolate original extraction's concentration for cases where we had to measure a dilution on the qubit

# convert the "dilution factor" into a proportion (original extract volume)/(H2O volume)
proportion = c() # put proportion values in a vector, in matching order with the data frame
for (i in 1:length(in_range_qs$extraction_id)) {
  dilfac = in_range_qs$dilution_factor[i]
  if (dilfac == "original_extract") {
    proportion[i] = 1
  } else if (dilfac == "1:1") {
    proportion[i] = 0.5
  } else if (dilfac == "1:3") {
    proportion[i] = 0.25
  }
}

qs_ng_ul = mutate(qs_ng_ul, proportion_original = proportion) # make that vector a column in qs_ng_ul


concs = mutate (qs_ng_ul, extrapolated_original_ng_per_ul = (1/proportion_original)*ng_per_ul) # calculate original concentration from measured concentration of dilution

head(concs)
```




```{r}
concs = mutate(concs, remaining_original_ul = ul_elution_original - ul_original_consumed - 10) # make new column with remaining ul of original extract, -10ul for safety/EB that never made it out of the spin column

concs$remaining_original_ul # looks like that worked
```
```{r}
library(ggplot2)

concs4graph = concs %>% mutate(date_extracted = as.character(date_extracted))
ggplot(concs4graph, aes(x = date_extracted, y = extrapolated_original_ng_per_ul)) +
  geom_boxplot() + geom_jitter()
```



# Now, we see what we need to do to dilute each extract to 2ng/ul 

c (concentration) = m (mass) / v (volume)

current state (1): we have a finite mass of DNA, in elution buffer, at a measured concentration (ng/uL)
    # c1 = qubit-measured concentration in ng/ul ("extrapolated_original_ng_per_ul")
    # m1 = total mass of DNA in usable volume of original extract
    # v1 = usable volume of original extract ("remaining_original_ul")
state 2: we want to use all the DNA in that usable volume of extract to make a 2ng/ul solution of unknown volume
    # c2 = 2ng/ul
    # m2 = m1; total mass of DNA in diluted extract (we're using all of the usable original)
    # v2 = total volume of diluted extract
* we need to find v2! and v2 will be made of v1+H2O, so volume of H2O needed will be v2-v1


c = m/v, so cv = m
m2 = m2, so c1v1 = c2v2, so (c1v1)/c2 = v2
and H2O needed = v2-v1

```{r}
# calculate final volume if we dilute all original extract
v2 = function(c1, v1, c2) {
  v2 = (c1*v1)/c2
  return(v2)
}

concs = mutate(concs, # add 2 columns: final volume of diluted extract, volume of H2O to add
               final_vol_ul = v2(extrapolated_original_ng_per_ul, remaining_original_ul, 2),
               H2O_ul = final_vol_ul - remaining_original_ul)
head(concs)


```


# # Dilutions for strip tubes (final vol 190ul)
```{r}
# we want the final dilution to fit in a strip tube (max v2 possible <190ul)

# so we constrain v2 and find v1; (c2v2)/c1 = v2
v1 = function(c2, v2, c1) {
  v1 = (c2*v2)/c1
  return(v1)
}

already_good = filter(concs, concs$final_vol_ul <= 190) # set aside cases where diluting all original extract would fit into a strip tube 
head(already_good)

concs200 = concs # make a copy of our data table to work in

diluting_whole_original = which(!(concs$extraction_id %in% (already_good$extraction_id))) # find indicies of extractions a dilution of the whole original volume will fit in a 200ul tube
diluting_aliquot = which(!(concs$extraction_id %in% (already_good$extraction_id))) # find indicies of extractions where we'd to dilute an aliquot to fit into 200ul strip tubes

concs200 = concs200 %>% 
  mutate(final_vol_ul = case_when(
    concs$extraction_id %in% (already_good$extraction_id) ~ final_vol_ul, # for extractions for which a dilution of the whole original volume will fit in a 200ul tube, set final vol of dilution as the final volume we find if we dilute the whole original extract
    !(concs$extraction_id %in% (already_good$extraction_id)) ~ 190)) # for extractions where we'd to dilute an aliquot to fit into 200ul strip tubes, set final vol of dilution to 190ul (200ul - 10ul for headspace)


# concs200[diluting_whole_original, "final_vol_ul"] = final_vol_if_all_diluted # for these, set final vol of dilution as the final volume we find if we dilute the whole original extract
# concs200[diluting_aliquot, "final_vol_ul"] = 190 # for these, set final vol of dilution to 190ul. 200ul - 10ul for headspace


# for extractions with these indices, find volume of original extraction we need to make a 2ng/ul dilution the final volume set above
concs200 = concs200 %>%
  mutate(ul_original_into_final = v1(2.01, # a little more than 2, so rounding errors don't push the final concentrations below 2
                                              final_vol_ul, # final vol set above
                                              extrapolated_original_ng_per_ul) # original extract concentrations
  ) %>%
  mutate(concs200, # re-write h2o vol
               H2O_ul = final_vol_ul - ul_original_into_final)

head(concs200)
```
```{r}

concs200 = concs200 %>% 
  
  # round volumes I'll be pipetting to the tenth place so I can do it with a p200
  mutate( 
    round_original = round(ul_original_into_final, digits = 1),
    round_H2O = round(H2O_ul, digits = 1),
    final_vol_ul = round_H2O + round_original
  ) %>%
  
  # calculate actual final concentration 
  mutate(final_ng_per_ul = c2(extrapolated_original_ng_per_ul, round_original, final_vol_ul))


concs200$final_ng_per_ul # good, once I re-set the final concentration in the code chunk above to 2.01

# code to check for rounding errors: 
head(filter(concs200, final_ng_per_ul < 2)) # if final conc is set anywhere 2.01 - 2.05, 7 show up as <2ng/ul final concentration actual
```

```{r}
# tidy 
DNA_dilutions_all_200 = concs200 %>% 
  select(!c(qubit_run_id, qubit_run_id_local, ng_per_ml, dilution_factor, proportion_original, ng_per_ul)) %>% # remove columns which were only relevant as metadata for qubit measurements
  rename(original_ng_per_ul = extrapolated_original_ng_per_ul, # rename concentration of original extract for clarity
         ul_H2O_into_final = H2O_ul,
         ul_original_used_for_qubit = ul_original_consumed) # 

DNA_dilutions_all_200 = DNA_dilutions_all_200[order(DNA_dilutions_all_200$extraction_id),] # reorder dataframe by extraction id

```

```{r}
write.csv(DNA_dilutions_all_200, "Results/DNA_dilutions_all_striptubes.csv")
```

